---
title: "grassEndophyteVignette"
author: "Robin Snyder"
date: "May 11, 2026"
output: pdf_document
bibliography: grassEndo.bib
link-citations: true
---


**Study background** This vignette involves perennial grasses with and
without endophytes.  We use the size-structured matrix model for
*Elymus villosus* from @2024:fowle.ziegl.whitn.ea, which includes
transition matrices for 13 years.  The model is in turn based on data
from a long-term symbiont removal experiment at Lilly-Dickey Woods in
Indiana, USA [@2023:mille.fowle.ziegl.ea].

**Variables**:

* PlistAllEndo: a list of lists containing the yearly transition
matrices: PlistAllEndo[[1]] is a list of yearly transition matrices
for plants with endophytes and PlistAllEndo[[2]] is the same for
plants without endophytes

* FlistAllEndo: a list of lists containing the yearly fecundity
matrices, with the same structure as PlistAllEndo

* Q: the transition matrix between year types.  (Here we assume that year
types are drawn randomly with replacement.)

* c0: The initial size distribution.  (Here all individuals are born
  into size class 1, the seedling size class, so c0[1] = 1 and all
  other entries are 0.)

* maxClutchSize: the maximum number of offspring that can be produced
  in a single reproductive bout

* maxLRO: the maximum lifetime reproductive output

* traitDist: the probability distribution for trait values.  Here we
  assume that plants are equally likely to have endophytes or not, so
  traitDist = c(0.5, 0.5).

* Fdist: the probability distribution for the clutch size, with
  current options "Poisson" and "Bernoulli".  Here we assume
  "Poisson". 

One question we may wish to ask is how certain we are of an
individual's endophyte status based on their lifetime reproductive
output (LRO).  In order to make the code less glacially slow,
we use an abridged version of the model, with only three year
types.  (WARNING: this code is still slow.)  For this species, not
having endophytes produces higher LRO, but even if we observe an
individual with 6 offspring (which is well into the 90th percentile),
we can't be very confident that that individual lacked endophytes.
For more details on the math behind this calculation, see
@2018:snyde.ellne and @2024:snyde.ellne.

```{r, eval=FALSE}
library ("luckieR")
maxClutchSize = 24
maxLRO = 24
groupDist = rep (0.5, 2)

## read in matrices
PlistAllEndo = readRDS ("ELRIPMatrices.rds")
FlistAllEndo = readRDS ("ELRIFMatrices.rds")

## find number of year types
numEnv = length(PlistAllEndo[[1]])

## Assume that all env. are equally likely and the sequence is
## temporally uncorrelated.  You can make other choices if you have
## more information.
Q = matrix (1/numEnv, numEnv, numEnv)

## How many states do we have in each env.?
mz = dim(PlistAllEndo[[1]][[1]])[1]

bigmz = mz * numEnv

## Initial state distrib.
c0 = rep (0, mz); c0[1] = 1

## Clutch size distribution
Fdist = "Poisson"

## Equally likely to have endophytes or not
traitDist = rep (0.5, 2)

out = probTraitCondLRO (PlistAllEndo, FlistAllEndo, Q,
                        c0, maxClutchSize, maxLRO,
                        traitDist, Fdist)

probXCondR = out$probXCondR

matplot (0:6, probXCondR[1:7,], xlab="LRO",
         ylab="Probability", type="l", col=c("black", "red"), lwd=2,
         lty=1, main="ELRI", ylim=c(0,1))

legend (x="topleft", col=c("black", "red"),
        legend=c("With endophytes", "Without endophytes"),
        lwd=2, bty="n", cex=1.3)
```

Alternatively, we may wish to focus on what forms of luck are most
important and at what ages.  If we partition the variance in LRO into
a within-endophyte-status contribution and a between-endophyte-status
contribution, we can think of the within-endophyte-status contribution
as the contribution of luck [@2018:snyde.ellne].  We can further
partition that luck contribution into contributions from individual
demographic transitions at different ages.  Did you survive from age 2
to age 3 and how much does knowing the outcome reduce the overall
variance?  If you lived, how much did you grow or shrink and how much
does knowing the outcome reduce the overall variance?  And so forth.
The amount that knowing the outcome of each of these demographic
transitions reduces the overall variance is the contribution that
transition makes.

We see that early survival luck is key and this makes sense: if you
die young, it doesn't matter how lucky you would have been later.  We
also see that growing quickly at ages 2 and 3 is important.  Survival
rises rapidly with size for this species, so this too, makes sense.
(Note: in order to make the code slow, we use an abridged version of
the model, with only three year types.)

```{r, eval=FALSE}
library ("luckieR")
maxClutchSize = 24
maxLRO = 24
groupDist = rep (0.5, 2)

## read in matrices
PlistAllEndo = readRDS ("ELRIPMatrices.rds")
FlistAllEndo = readRDS ("ELRIFMatrices.rds")

## find number of year types
numEnv = length(PlistAllEndo[[1]])

## Assume that all env. are equally likely and the sequence is
## temporally uncorrelated.  You can make other choices if you have
## more information.
Q = matrix (1/numEnv, numEnv, numEnv)

## How many states do we have in each env.?
mz = dim(PlistAllEndo[[1]][[1]])[1]

bigmz = mz * numEnv

## Initial state distrib.
c0 = rep (0, mz); c0[1] = 1
## Initial env. distrib.  Should probably be the stationary distribution.
u0 = rep (1/numEnv, numEnv)

## We have two traits: endophyte-positive and -negative
numTraits=2
## Maximum age
maxAge = 25

## Equally likely to have endophytes or not
traitDist = rep (0.5, 2)

out = partitionVarSkewnessEnvVarAndTraits (PlistAllEndo,
                                           FlistAllEndo,
                                           Q, c0, traitDist, maxAge,
                                           survThreshold=0.05)

varTrajecResults = array (NA, dim=c(numTraits, 4, maxAge))
varTrajecResults[,1,] = out$survTrajecVarCondX
varTrajecResults[,2,] = out$growthTrajecVarCondX
varTrajecResults[,3,] = out$envTrajecVarCondX
varTrajecResults[,4,] = out$fecVarCondX

colorVec = colorVecAll6[2:5]
matplot (1:9, t(varTrajecResults[1,,1:9]),
         xlab="Age", ylab="Contrib. to Var(R)", type="l",
         main="ELRI with endophytes", lwd=2,
	 col=c("black", "orange", "forest green", "turquoise"),
	 lty=1, cex.lab=1.3)

legend (x="topright", lwd=2,
        col=c("black", "orange", "forest green", "turquoise"),
        legend=c("Survival", "Growth", "Environment", "Fecundity"),
        lty=1, cex=1.4, bty="n")
```